---
title: "Data cleaning Pythium"
author: "Daniel Cerritos"
date: "2/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(readxl)
library(tidyverse)
library(here)
library(naniar)
```

Pythium species were incorreclty identified during BLAST search, so the NK lab had wrong species.  
NK's lab used a plate number id for the assays, so I would combine correct identification with the ids from assays.

```{r pythium ids}
# our update lab list
pythium_ids <- read_excel(here("data", "Pythium ID 2016-2017 (1).xlsx"), skip = 2,
                          col_names = c("sample_id", "county_sample", "gps_N", 
                                        "gps_W", "isolate_stock", "county_isolate", 
                                        "database_binomial", "amplified_by", 
                                        "primers", "trace_file_name", "length_bp", 
                                        "ncbi_blast_taxonomy", "blast_score", 
                                        "blast_clade", "philo_species", 
                                        "philo_accession", "philo_similarity", 
                                        "philo_note"))

# not updated list nathan's lab had
isolates_used <- read_excel(here("data", "Final Pythium survey isolate IDs 2018.xlsx"), 
                            col_names = c("plate_id", "isolate_id", "species",
                                          "date_sequenced", "notes"), skip = 1) %>% 
  dplyr::select(plate_id, isolate_id)

# combine both lists
correct_ids <- pythium_ids %>% 
  dplyr::select(sample_id, county = county_sample, isolate_id = isolate_stock, 
                species = philo_species)

plate_and_correct_ids <- right_join(correct_ids, isolates_used, by = "isolate_id") %>% 
  arrange(plate_id)

# save
# write_csv(plate_and_correct_ids, path = here("output", "clean_data", "01_list-of-isolates-used_clean-data.csv"))
```

## Soybean and maize seed aggressiveness 

### Soybean

```{r seed soybean}

seed_soybean <- read_excel(here("data", "Pythium isolate survey 2018 aggressiveness- soybean.xlsx"), 
                           sheet = 2, skip = 4)

# select what you need
seed_soybean <- seed_soybean %>% 
  dplyr::select(Species, plate_id = id, rep, colonized_5 = fuzzy, colonized_10 = ...5, 
                rotted, germination)

# seperate controls because they are characters
seed_soybean_controls <- seed_soybean %>% 
  filter(plate_id %in% c("pda", "pi", "pu")) %>% 
  mutate(
    isolate_id = "control",
    species = case_when(
    plate_id == "pda" ~ "PDA control", 
    plate_id == "pi" ~ "Pu control", 
    plate_id == "pu" ~ "Pi control"
  ))

seed_soybean_isolates <- seed_soybean %>% 
  filter(!plate_id %in% c("pda", "pi", "pu"))

# ids and plates 
add_names <- prueba %>% 
  dplyr::select(isolate_id, species, plate_id)

# combine
seed_soybean_isolates$plate_id <- as.numeric(seed_soybean_isolates$plate_id)
seed_soybean_isolates_correct <- right_join(add_names, seed_soybean_isolates, by = "plate_id")
seed_soybean_isolates_correct$plate_id <- as.character(prueba_2$plate_id)

# all data
seed_soybean_clean <- bind_rows(seed_soybean_controls, seed_soybean_isolates_correct)
# turn "-" into NAs
seed_soybean_clean <- seed_soybean_clean %>% 
    replace_with_na_all(condition = ~.x == "-")

# save
write_csv(seed_soybean_clean, path = here("output", "clean_data", "01_seed-aggressiveness-soybean_clean-data.csv"))
```

### Corn

```{r seed corn}
seed_corn <- read_excel(here("data", "Pythium isolate survey Aggressiveness -corn.xlsx"), 
                        sheet = 2, skip = 3)

seed_corn <- seed_corn %>% 
  dplyr::select(Species, plate_id = id, run, rep = Rep, colonized = fuzzy, germination, 
                bag_weight_g = `Bag weight`, dry_weight_bag_g =`Dry Mass(w/ bag)`, 
                )

# seperate controls and isolate data
seed_corn_isolates <- seed_corn %>% 
  filter(!plate_id %in% c("pda", "control"))

seed_corn_isolates$plate_id <- as.numeric(seed_corn_isolates$plate_id)
seed_corn_isolates_names <- right_join(add_names, seed_corn_isolates, by = "plate_id")

seed_corn_controls <- seed_corn %>% 
  filter(plate_id %in% c("pda", "control")) %>% 
  mutate(isolate_id = "control", 
         species = case_when(
           plate_id == "pda" ~ "PDA control", 
           plate_id == "control" ~ "Pythium control", 
           plate_id == "control" ~ "Pythium control"
         ))

# combine both
seed_corn_isolates_names$plate_id <- as.character(seed_corn_isolates_names$plate_id)
seed_corn_clean <- bind_rows(seed_corn_controls, seed_corn_isolates_names)

# save
write_csv(seed_corn_clean, path = here("output", "clean_data", "01_seed-aggressiveness-corn_clean-data.csv"))
```


## Soybean seedling aggressivenes 

### agar and soil inocolum 

```{r plug pot assay }
soybean_plug <- read_excel(here("data", "pot assay.xlsx"), sheet = 2, skip = 2)
soybean_plug <- soybean_plug %>% 
  dplyr::select(plate_id =  `Plate ID`, rep = Rep, isolate_id = `Isolate IDs`,
                emergence = `Emerged?`, plant_mass_g = `Plant mass`, root_mass_g = `Root mass`, 
                root_rot_perc = `% root rot`)

# remove g from observation 
soybean_plug$plant_mass_g <- as.double(gsub('[a-zA-Z]', '', soybean_plug$plant_mass_g))
soybean_plug$root_mass_g <- as.double(gsub('[a-zA-Z]', '', soybean_plug$root_mass_g))

# combine 
soybean_plug_clean <- right_join(add_names, soybean_plug)
soybean_plug_clean <- soybean_plug_clean %>% 
  replace_with_na_all(condition = ~.x == "-")
# save
write_csv(soybean_plug_clean, path = here("output", "clean_data", "01_seedling-aggressiveness-soybean-agar_clean-data.csv"))
```

### millet inocolum

```{r millet assay }
# millet  
soybean_millet <- read_excel(here("data", "Pythium isolate aggressiveness-pot experiment.xlsx"), 
                             sheet = 2, skip = 2)

soybean_millet <- soybean_millet %>% 
  rename(total_plant = `total plants`, germination_per = `germination (%)`, 
         envelope_g = `envelope mass (g)`, envelope_plant_g = `envelope + plant mass (g)`, 
         dry_weight = `dry mass (g)`) %>% 
  dplyr::select(-`dry mass average (g)`)

soybean_millet <- soybean_millet %>% 
  mutate(
    millet_ml = case_when(
      millet %in% c("10 ml sterile (control)", "10 ml Pu", "10 ml Pi", "10 ml Ps") ~ 10, 
      millet %in% c("20 ml sterile (control)", "20 ml Pu", "20 ml Pi", "20 ml Ps") ~ 20,
      millet == "None" ~ 0))
    
soybean_millet <- soybean_millet %>% 
  mutate(
    species = case_when(
      millet == "None" ~ "control",
      millet %in% c("10 ml sterile (control)", "20 ml sterile (control)") ~ "millet control", 
      millet %in% c("10 ml Pu", "20 ml Pu") ~ "Pythium ultimum var. ultimum" , 
      millet %in% c("10 ml Pi", "20 ml Pi") ~ "Pythium irregulare", 
      millet %in% c("10 ml Ps", "20 ml Ps") ~ "Pythium sylvaticum"
      )
        )

write_csv(soybean_millet, path = here("output", "clean_data", "01_seedling-aggressiveness-soybean-millet_clean-data.csv"))
```

## Effect of media on pythium and soybean growth 


```{r media soybean}
media_soybean <- read_excel(here("data", "aggressiveness-media.xlsx"), 
                             sheet = 2, skip = 4)

media_soybean <- media_soybean %>% 
  dplyr::select(media = Media, innocolum = Innoculm, run, colonized_5 = fuzzy...5, 
                germination_5 = germination...6, rotted_5 = rotted...7,
                colonized_10 = fuzzy...8, germination_10 = germination...9, 
                rotted_10 = rotted...10)

# replace "-" with NAs
media_soybean <- media_soybean %>% 
  replace_with_na_all(condition = ~.x == "-")
```

## fungicide sensitivity 

```{r fs pythium }

# import files 
azoxystrobin <- read_excel(here("data", "Pythium_Azoxystrobin.xlsx")) %>% 
  mutate(fungicide = "Azoxystrobin")

ethaboxam <- read_excel(here("data", "Pythium_Ethaboxam.xlsx")) %>% 
  mutate(fungicide = "Ethaboxam")

metalaxyl <- read_excel(here("data", "Pythium_Metalaxyl.xlsx")) %>% 
  mutate(fungicide = "Metalaxyl")

mefenoxam <- read_excel(here("data", "Pythium_Mefenoxam.xlsx")) %>% 
  mutate(fungicide = "Mefenoxam")

# combine in one file
all_fungicide_data <- bind_rows(azoxystrobin, ethaboxam, metalaxyl, mefenoxam)

write_csv(all_fungicide_data, here("output", "clean_data", "01-fungicide-sensitivty_clean-data.csv"))
```
