---
title: "Exploratory analysis Pythium clean data"
author: "Daniel Cerritos"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(here)
library(tidyverse)
theme_set(theme_light())
```

```{r import data}
# aggressiveness seed assays
seed_soybean <- read_csv(here("output", "clean_data", "01_seed-aggressiveness-soybean_clean-data.csv"))
seed_corn <- read_csv(here("output", "clean_data", "01_seed-aggressiveness-corn_clean-data.csv"))
# aggressiveness seedling assays
seedling_soybean_agar <- read_csv(here("output", "clean_data", "01_seedling-aggressiveness-soybean-agar_clean-data.csv"))
seedling_soybean_millet <- read_csv(here("output", "clean_data", "01_seedling-aggressiveness-soybean-millet_clean-data.csv"))
# effect of media on pythium growth 
media_soybean <- read_csv(here("output", "clean_data", "01_media-effect-pythium-aggressiveness.csv"))
# fungicide sensitivity
fungicide <- read_csv(here("output", "clean_data", "01-fungicide-sensitivty_clean-data.csv"))

```


## Aggressiveness on seed

### Soybean

```{r}
seed_soybean %>% 
  ggplot(aes(x = reorder(rep, rotted_5), y = rotted_5)) +
  geom_boxplot() + 
  geom_jitter() 
  facet_wrap(vars(rep)) +
  coord_flip()
```





```{r}
# check controls across reps
seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, rotted_5), y = rotted_5)) +
  geom_boxplot() + 
  geom_jitter() +
  facet_wrap(vars(rep)) +
  coord_flip()

seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5, color =  species)) +
  geom_jitter() +
  facet_wrap(vars(rep)) 

seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, colonized_10), y = colonized_10, color =  species)) +
  geom_jitter() +
  facet_wrap(vars(rep)) 

seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, germination_10), y = germination_10, color =  species)) +
  geom_jitter() +
  facet_wrap(vars(rep)) 
```

Controls look fine, the responses are similar across different set of experiments 

Differences b/w species?

```{r soy seed agg species}
seed_soybean %>% 
  filter(!is.na(germination_10)) %>% 
  ggplot(aes(x = reorder(species, germination_10), y = germination_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 

seed_soybean %>% 
  filter(!is.na(colonized_5)) %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip()

seed_soybean %>% 
  filter(!is.na(colonized_10)) %>% 
  ggplot(aes(x = reorder(species, colonized_10), y = colonized_10))+
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() # more differences observed at 10 days

seed_soybean %>% 
  filter(!is.na(rotted_5)) %>% 
  ggplot(aes(x = reorder(species, rotted_5), y = rotted_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() # not that informative, probably drop variable
```

Difference within species?

```{r within species seed soybean}
# focus on day 10 observations
# filter the most aggressive species, no difference it seems
most_agg_species <- c("Pythium ultimum var. ultimum", "Pythium ultimum var. sporangiiferum", 
                       "Pythium aphanidermatum", "Pu control", "Pi control", "Pythium irregulare")

seed_soybean %>% 
  filter(!is.na(colonized_5), 
         species %in% most_agg_species) %>% 
  ggplot(aes(x = reorder(plate_id, colonized_5), y = colonized_5, color = species)) +
  geom_point(stat = "summary") +
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width = 0.2) # no difference at 5 days

seed_soybean %>% 
  filter(!is.na(germination_10), 
         !species %in% most_agg_species) %>% 
  ggplot(aes(x = reorder(plate_id, germination_10), y = germination_10, color = species)) +
  geom_point(stat = "summary") +
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width = 0.2)


seed_soybean %>% 
  filter(!is.na(colonized_10), 
         !species %in% most_agg_species) %>% 
  ggplot(aes(x = reorder(plate_id, colonized_10), y = colonized_10, color = species)) +
  geom_point(stat = "summary") 
```

it looks there is some variation b/w isolates of same species, but for more aggressive species doesn't seem to be variation

### Corn 


```{r check controls}
seed_corn %>% 
  filter(isolate_id == "control") %>% 
  ggplot(aes(x = reorder(species, germination_5), y = germination_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(run))

seed_soybean %>% 
  filter(isolate_id == "control") %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(rep))

seed_corn %>% 
  mutate(dry_weight_g = dry_weight_bag_g - bag_weight_g) %>% 
  filter(isolate_id == "control") %>% 
  ggplot(aes(x = reorder(species, dry_weight_g), y = dry_weight_g)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(rep))
```

Controls across the different sets and runs look the same.


```{r}
# distributions
seed_corn %>% 
  ggplot(aes(y = germination_5)) +
  geom_histogram()
  
seed_corn %>% 
  ggplot(aes(y = colonized_5)) +
  geom_histogram()
  
seed_corn %>% 
  mutate(dry_weight_g = dry_weight_bag_g - bag_weight_g) %>% 
  ggplot(aes(y = dry_weight_g)) +
  geom_histogram()

```

Differences b/w species?

```{r corn seed agg species}
seed_corn %>%
  filter(!species == "Pi control") %>% 
  ggplot(aes(x = reorder(species, germination_5), y = germination_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip()

seed_corn %>% 
  filter(!species == "Pi control") %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip()

seed_corn %>% 
  filter(!species == "Pi control") %>% 
  mutate(dry_weight_g = dry_weight_bag_g - bag_weight_g) %>%  
  filter(!is.na(dry_weight_g)) %>% 
  ggplot(aes(x = reorder(species, dry_weight_g), y = dry_weight_g)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip()

# missing: yorkensis, irregulare, acrogynum, acanthophoron
```

Differences within species?

```{r}
seed_corn %>% 
  filter(!is.na(germination_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(plate_id, germination_5), y = germination_5, color = species)) +
  geom_point(stat = "summary") +
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width = 0.2)


seed_corn %>% 
  filter(!is.na(colonized_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(plate_id, colonized_5), y = colonized_5, color = species)) +
  geom_point(stat = "summary") +
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width = 0.2)

seed_corn %>% 
  mutate(dry_weight_g = dry_weight_bag_g - bag_weight_g) %>% 
  filter(!is.na(dry_weight_g), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(plate_id, dry_weight_g), y = dry_weight_g, color = species)) +
  geom_point(stat = "summary") +
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width = 0.2)
```

## Aggressiveness seedling assays

### Agar
```{r}
seedling_soybean_agar %>% 
  filter(!is.na(species)) %>% 
  mutate(emergence_bi = case_when(
    emergence == "N" ~ 0, 
    emergence == "Y" ~ 1,
    )) %>% 
  filter(!is.na(emergence_bi)) %>% 
  ggplot(aes(x = reorder(species, emergence_bi), y = emergence_bi)) +
  geom_col() +
  coord_flip() # no idea how to manage this data

seedling_soybean_agar %>%
  filter(!is.na(species), 
         !is.na(plant_mass_g)) %>% 
  ggplot(aes(x = reorder(species, plant_mass_g), y = plant_mass_g)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip()

seedling_soybean_agar %>%
   filter(!is.na(species), 
         !is.na(root_mass_g)) %>% 
  ggplot(aes(x = reorder(species, root_mass_g), y = root_mass_g)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip()
```

### Millet

## Effect of media on pythium growth 


```{r}
media_soybean %>% 
  ggplot(aes(x = reorder(type, germination_5), y = germination_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(innocolum), scales = "free_x")

media_soybean %>% 
  ggplot(aes(x = reorder(type, germination_10), y = germination_10)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(innocolum), scales = "free_x")


media_soybean %>% 
  ggplot(aes(x = reorder(type, colonized_5), y = colonized_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(innocolum), scales = "free_x")
  
media_soybean %>% 
  ggplot(aes(x = reorder(type, colonized_10), y = colonized_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(innocolum), scales = "free_x") # not much difference from day 5

# not using the rotted data, most is missing data
media_soybean %>% 
  ggplot(aes(x = reorder(type, rotted_5), y = rotted_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(innocolum), scales = "free_x")
  
media_soybean %>% 
  ggplot(aes(x = reorder(type, rotted_10), y = rotted_5)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(vars(innocolum), scales = "free_x") 
```

