---
title: "Analysis Pythium aggressiveness"
author: "Daniel Cerritos"
date: "2/2/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, message=FALSE}
library(here)
library(tidyverse)
library(scales)
library(patchwork)
library(cowplot)
library(ggfortify)
library(ggResidpanel)
library(lmerTest)
library(emmeans)
library(drc)
library(patchwork)
theme_set(theme_light())
```

```{r import data, message=FALSE}
# recovered isolates 
isolates_recovered <- read_csv(here("data", "all-isolates-in-data-base.csv"))
# aggressiveness seed assays
seed_soybean <- read_csv(here("output", "clean_data", "01_seed-aggressiveness-soybean_clean-data.csv"))
seed_corn <- read_csv(here("output", "clean_data", "01_seed-aggressiveness-corn_clean-data.csv"))
# aggressiveness seedling assays
seedling_soybean_agar <- read_csv(here("output", "clean_data", "01_seedling-aggressiveness-soybean-agar_clean-data.csv"))
seedling_soybean_millet <- read_csv(here("output", "clean_data", "01_seedling-aggressiveness-soybean-millet_clean-data.csv"))
# effect of media on pythium growth 
media_soybean <- read_csv(here("output", "clean_data", "01_media-effect-pythium-aggressiveness.csv"))
# fungicide sensitivity
fungicide <- read_csv(here("output", "clean_data", "01-fungicide-sensitivty_clean-data.csv"))

```

```{r}
# recreate map you use in the poster?
# just use the isolates used in the experiments
isolates_species <- isolates_recovered %>% 
  dplyr::select(species) %>%
  group_by(species) %>% 
  count() %>%  
  mutate(percentage = (n/149))

isolates_species %>% 
  filter(!species  %in% c("Phytophthora sansomeana", "Phytophthora sojae", "Phytophthora spp." )) %>% 
  ggplot(aes(x = reorder(species, percentage), y = percentage)) +
  geom_col(alpha = 0.5) +
  coord_flip() +
  labs(y = "Isolates recovered", 
       x = "") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     labels = percent_format()) +
  theme(legend.position = "none",
        axis.text.y  = element_text(face = "italic")) +
  scale_fill_manual(values = c("#BEBEBE"))

```

# Soybean Seed aggressiveness 

## Experiments 

Check variance across experiments

```{r}
seed_soybean %>% 
  filter(!is.na(rotted_5)) %>% 
  ggplot(aes(x = factor(experiment), y = rotted_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) 

seed_soybean %>% 
  filter(!is.na(colonized_5)) %>% 
  ggplot(aes(x = factor(experiment), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) # similar variance 

seed_soybean %>% 
  filter(!is.na(colonized_10)) %>% 
  ggplot(aes(x = factor(experiment), y = colonized_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) # similar variance 

seed_soybean %>% 
  filter(!is.na(germination_10)) %>% 
  ggplot(aes(x = factor(experiment), y = germination_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) # similar variance 
```

```{r}
seed_soybean %>% 
  filter(!is.na(rotted_5)) %>% 
  ggplot(aes(x = reorder(species, rotted_5), y = rotted_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 

seed_soybean %>% 
  filter(!is.na(colonized_5)) %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 

seed_soybean %>% 
  filter(!is.na(colonized_10)) %>% 
  ggplot(aes(x = reorder(species, colonized_10), y = colonized_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 

seed_soybean %>% 
  filter(!is.na(germination_10)) %>% 
  ggplot(aes(x = reorder(species, germination_10), y = germination_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 
```

## Controls

Reproducibility of assays, check controls across different reps

```{r}
# check controls across reps
seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, rotted_5), y = rotted_5)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(rep)) +
  coord_flip()

seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(rep)) +
  coord_flip()

seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, colonized_10), y = colonized_10)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(rep)) +
  coord_flip()

seed_soybean %>% 
  filter(isolate_id %in% c("PDA control", "Pu control", "Pi control")) %>% 
  ggplot(aes(x = reorder(species, germination_10), y = germination_10)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(rep)) +
  coord_flip()
```

## Fit models 


```{r}

seed_soybean %>% view()
seed_soybean$rep <- as.factor(seed_soybean$rep)
seed_soybean$isolate_id <- as.factor(seed_soybean$isolate_id)
seed_soybean$species <- as.factor(seed_soybean$species)
```

Include experiment as a factor

```{r fm_seed_agg_col_5}
fm_seed_agg_col_5 <- glm(colonized_5 ~ species + isolate_id + rep, data = seed_soybean)
qplot(fm_seed_agg_col_5$residuals)
resid_panel(fm_seed_agg_col_5)
anova(fm_seed_agg_col_5, test = "Chisq")
# mean separation
emmeans(fm_seed_agg_col_5, ~ species, type = "response") %>% 
    multcomp::cld()
emmeans(fm_seed_agg_col_5, ~ isolate_id|species, type = "response") %>% 
    multcomp::cld()
```

```{r fm_seed_agg_col_10}
fm_seed_agg_col_10 <- glm(colonized_10 ~ species + isolate_id + rep, data = seed_soybean)
resid_panel(fm_seed_agg_col_10)
anova(fm_seed_agg_col_10, test = "Chisq")
# mean separation
emmeans(fm_seed_agg_col_10, ~ species,  type = "response") %>% 
    multcomp::cld()
emmeans(fm_seed_agg_col_10, ~ isolate_id|species, type = "response") %>% 
    multcomp::cld()
```

```{r fm_seed_agg_ger_10}
fm_seed_agg_ger_10 <- lm(germination_10 ~ species + isolate_id + rep, data = seed_soybean)
qplot(fm_seed_agg_ger_10$residuals)
resid_panel(fm_seed_agg_ger_10)
anova(fm_seed_agg_ger_10, test = "Chisq")
# mean separation 
emmeans(fm_seed_agg_ger_10, ~ species, type = "response") %>% 
  multcomp::cld()
emmeans(fm_seed_agg_ger_10, ~ isolate_id|species, type = "response" ) %>% 
  multcomp::cld()
```


## Figures

```{r}
# reduce the binomials length, for plotting
seed_soybean <- seed_soybean %>% 
  mutate(short_species = case_when(
    species == "Pythium acrogynum" ~ "P. acrogynum", 
    species == "Pythium irregulare" ~ "P. irregulare", 
    species == "Pythium torulosum" ~ "P. torulosum",
    species == "Pythium ultimum var. ultimum" ~ "P. ultimum var. ultimum", 
    species == "Pythium yorkensis"  ~ "P. yorkensis", 
    species == "Pythium acanthophoron" ~ "P. acanthophoron", 
    species ==  "Pythium aphanidermatum" ~  "P. aphanidermatum", 
    species == "Pythium pleroticum"  ~ "P. pleroticum", 
    species == "Pythium ultimum var. sporangiiferum" ~ "P. ultimum var. sporangiiferum", 
    species == "Pythium vexans" ~ "P. vexans" , 
    TRUE ~ as.character(species)
    ))
```

```{r fig.height=10, fig.width=5}
soy_col_5 <- seed_soybean %>% 
  filter(!is.na(colonized_5)) %>% 
  ggplot(aes(x = reorder(short_species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  labs(y = "Colonized (5 dai)", 
       x = "")

soy_col_10 <-seed_soybean %>% 
  filter(!is.na(colonized_10)) %>% 
  ggplot(aes(x = reorder(short_species, colonized_10), y = colonized_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  labs(y = "Colonized (10 dai)", 
       x = "")

# combine colonized 
soybean_colonized <- soy_col_5 / soy_col_10 + plot_annotation(tag_levels = 'A')

ggsave("figure1.tiff", soybean_colonized, dpi = 300, 
       path = here("docs", "figures"), width = 5, height = 7)


soybean_germination <- seed_soybean %>% 
  filter(!is.na(germination_10)) %>% 
  ggplot(aes(x = reorder(short_species, germination_10), y = germination_10)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  labs(y = "Germination (10 dai)", 
       x = "")

ggsave("figure2.tiff", soybean_germination, dpi = 300, 
       path = here("docs", "figures"), width = 5, height = 5)

```

# Corn Seed aggressiveness 

```{r}
# subtract paper bag weight 
seed_corn <- seed_corn %>% 
  mutate(weight_g = dry_weight_bag_g - bag_weight_g)
```

## Controls

Check controls across sets 

```{r}
# check controls across reps
# drop Pi control, only in one set
seed_corn %>% 
  filter(isolate_id %in% c("PDA control", "Pu control")) %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(run)) 

seed_corn %>% 
  filter(isolate_id %in% c("PDA control", "Pu control")) %>% 
  ggplot(aes(x = reorder(species, germination_5), y = germination_5)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(run)) 

seed_corn %>% 
  filter(isolate_id %in% c("PDA control", "Pu control")) %>% 
  ggplot(aes(x = reorder(species, weight_g), y = weight_g)) +
  geom_boxplot() + 
  geom_jitter(aes(color = species), alpha = 0.4) +
  facet_wrap(vars(run)) 
```


## Experiment

Check variances across the two experiment 


```{r}
seed_corn %>% 
  filter(!is.na(colonized_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 
# isolates 8 and 10 there first and second run were in 3 and 4

seed_corn %>% 
  filter(!is.na(colonized_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(species, germination_5), y = germination_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 

seed_corn %>% 
  filter(!is.na(colonized_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(species, weight_g), y = weight_g)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  facet_wrap(vars(experiment), scales = "free_x") 
```

## Fit models

```{r}
seed_corn$run <- as.factor(seed_corn$run)
seed_corn$isolate_id <- as.factor(seed_corn$isolate_id)
seed_corn$species <- as.factor(seed_corn$species)
seed_corn <- seed_corn %>% 
  filter(!species == "Pi control")

```

```{r}
fm_seed_corn_col_5 <- lm(colonized_5 ~ species + isolate_id + run, data = seed_corn)
qplot(fm_seed_corn_col_5$residuals)
autoplot(fm_seed_corn_col_5, which = 1:4, ncol = 2, label.size = 3)
anova(fm_seed_corn_col_5) # run not significant  

# remove run
fm_seed_corn_col_5 <- glm(colonized_5 ~ species + isolate_id, data = seed_corn)
qplot(fm_seed_corn_col_5$residuals)
autoplot(fm_seed_corn_col_5, which = 1:4, ncol = 2, label.size = 3)
anova(fm_seed_corn_col_5, test = "Chisq")
emmeans(fm_seed_corn_col_5, ~ species) %>% 
  multcomp::cld()
```

```{r}
fm_seed_corn_ger_5 <- lm(germination_5 ~ species + isolate_id + run, data = seed_corn)
qplot(fm_seed_corn_ger_5$residuals)
autoplot(fm_seed_corn_ger_5, which = 1:6, ncol = 3, label.size = 3)
anova(fm_seed_corn_ger_5) # run not significant  

# remove run
fm_seed_corn_ger_5 <- glm(germination_5 ~ species + isolate_id, data = seed_corn)
qplot(fm_seed_corn_ger_5$residuals)
autoplot(fm_seed_corn_ger_5, which = 1:6, ncol = 3, label.size = 3)
anova(fm_seed_corn_ger_5, test = "Chisq") 
emmeans(fm_seed_corn_ger_5, ~ species) %>% 
  multcomp::cld()
emmeans(fm_seed_corn_ger_5, ~ isolate_id|species) %>% 
  multcomp::cld()
```

```{r}
# I think I will not use this data
fm_seed_corn_weig <- lm(weight_g ~ species + isolate_id + run, data = seed_corn)
qplot(fm_seed_corn_weig$residuals)
autoplot(fm_seed_corn_weig, which = 1:4, ncol = 2, label.size = 3)
anova(fm_seed_corn_weig) # not remove
emmeans(fm_seed_corn_weig, ~ species) %>% 
  multcomp::cld()

```


## Figures

```{r}
# shorter species 
seed_corn <- seed_corn %>% 
  mutate(short_species = case_when(
    species == "Pythium torulosum" ~ "P. torulosum",
    species == "Pythium ultimum var. ultimum" ~ "P. ultimum var. ultimum", 
    species ==  "Pythium aphanidermatum" ~  "P. aphanidermatum", 
    species == "Pythium pleroticum"  ~ "P. pleroticum", 
    species == "Pythium ultimum var. sporangiiferum" ~ "P. ultimum var. sporangiiferum", 
    species == "Pythium vexans" ~ "P. vexans" , 
    TRUE ~ as.character(species)
    ))
```

```{r}
corn_colonized <- seed_corn %>% 
  filter(!is.na(colonized_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(short_species, colonized_5), y = colonized_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  labs(y = "Colonized (5 dai)", 
       x = "")

corn_germination <- seed_corn %>% 
  filter(!is.na(germination_5), 
         !species == "Pi control") %>% 
  ggplot(aes(x = reorder(short_species, germination_5), y = germination_5)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  labs(y = "Germination (10 dai)", 
       x = "")

# combine colonized 
corn_both <- corn_colonized / corn_germination + plot_annotation(tag_levels = 'A')

ggsave("figure3.tiff", corn_both, dpi = 300, 
       path = here("docs", "figures"), width = 5, height = 6)


seed_corn %>% 
  filter(!is.na(weight_g)) %>% 
  ggplot(aes(x = reorder(short_species, weight_g), y = weight_g)) +
  geom_boxplot() + 
  geom_jitter (alpha = 0.2) +
  coord_flip() +
  labs(y = "Dry weight (g)", 
       x = "")
```


# Aggressiveness soybean seedling assays

## millet 

```{r}
seedling_soybean_millet <- seedling_soybean_millet %>%
  mutate(dry_weight_g = envelope_plant_g - envelope_g)

seedling_soybean_millet$run <- as.factor(seedling_soybean_millet$run)
seedling_soybean_millet$millet_ml <- as.factor(seedling_soybean_millet$millet_ml)
seedling_soybean_millet$species <- as.factor(seedling_soybean_millet$species)
```


```{r}
fm_millet_ger <- lm(germination ~ species + millet_ml + run, data = seedling_soybean_millet)
qplot(fm_millet_ger$residuals)
autoplot(fm_millet_ger, which = 1:4, ncol = 2, label.size = 3)
anova(fm_millet_ger) # no diff from millet and run

# remove millet and run from model 
fm_millet_ger <- lm(germination ~ species, data = seedling_soybean_millet)
qplot(fm_millet_ger$residuals)
autoplot(fm_millet_ger, which = 1:4, ncol = 2, label.size = 3)
anova(fm_millet_ger)
emmeans(fm_millet_ger, ~ species) %>% 
  multcomp::cld()
```

```{r}
fm_millet_weight <- lm(dry_weight_g ~ species + millet_ml + run, data = seedling_soybean_millet)
qplot(fm_millet_weight$residuals)
autoplot(fm_millet_weight, which = 1:4, ncol = 2, label.size = 3)
anova(fm_millet_weight) # no diff from millet and run

# remove millet and run from model 
fm_millet_weight <- lm(dry_weight_g ~ species, data = seedling_soybean_millet)
qplot(fm_millet_weight$residuals)
autoplot(fm_millet_weight, which = 1:4, ncol = 2, label.size = 3)
anova(fm_millet_weight)
emmeans(fm_millet_weight, ~ species) %>% 
  multcomp::cld()
```


## Figure

```{r}
seedling_soybean_millet <- seedling_soybean_millet%>% 
  mutate(short_species = case_when(
    species == "Pythium sylvaticum" ~ "P. sylvaticum",
    species == "Pythium ultimum var. ultimum" ~ "P. ultimum var. ultimum", 
    species ==  "Pythium irregulare"  ~  "P. irregulare", 
    TRUE ~ as.character(species)
    ))
```

```{r}
seedling_germination <- seedling_soybean_millet %>%
  ggplot(aes(x = reorder(short_species, germination), y = germination)) +
  geom_boxplot(alpha = 0.5) + 
  geom_jitter (alpha = 0.2, color = "gray50") +
  coord_flip() +
  labs(y = "Germination (10 dai)", 
       x = "")
  
seedling_weight <- seedling_soybean_millet %>%
  ggplot(aes(x = reorder(short_species, dry_weight_g ), y = dry_weight_g)) +
  geom_boxplot(alpha = 0.5) + 
  geom_jitter (alpha = 0.2, color = "gray50") +
  coord_flip() +
  labs(y = "Dry weight (g)", 
       x = "")

# combine colonized 
seedling_both <- seedling_germination / seedling_weight + plot_annotation(tag_levels = 'A')

ggsave("figure4.tiff", seedling_both, dpi = 300, 
       path = here("docs", "figures"), width = 4, height = 5)



```

# Fungicide sensitivity

```{r}
# eliminate NAs
# average colony measurements 
transform_fungicide <- fungicide %>% 
  filter(!is.na(a)) %>%  
  mutate(length_mm = (a + b)/2) 

# subtract the plug diameter
transform_fungicide <- transform_fungicide %>% 
  mutate(length_mm = case_when(
    length_mm > 0 ~ length_mm - 4, 
    TRUE ~ length_mm - 0
    )) %>% 
  filter(!length_mm < 0) # eliminate any negative values

# mean growth of control plate (dose 0)
control_fungicide <- transform_fungicide  %>% 
  filter(dose == 0, 
         !is.na(length_mm)) %>% 
  group_by(fungicide, run, isolate_id) %>% 
  summarise(control_mean = mean(length_mm)) 

transform_fungicide <- left_join(transform_fungicide, control_fungicide, by = c("fungicide",
                                                                                "run", 
                                                                                "isolate_id"))

# relative growth calculation
transform_fungicide  <- transform_fungicide  %>% 
  mutate(relative_growth = (length_mm/control_mean)*100)

# write_csv(transform_fungicide, here("output", "transform_data", "03_fungicide-sensitivity_transform.csv"))
```

```{r}
transform_fungicide %>% 
  group_by(species, isolate_id) %>% 
  tally()
```

```{r}
# check if there are isolates with no EC50
# combine both trials
mean_relative_growth <- transform_fungicide %>% 
group_by(fungicide, isolate_id, dose, species) %>% 
  summarise(relative_growth = mean(relative_growth))

# isolates with >50% rel growth in the higher dose, no EC50 can be calculated
# Azoxystrobin
mean_relative_growth %>% 
  filter(fungicide %in% c("Mefenoxam","Metalaxyl", "Azoxystrobin") & dose == 10 & relative_growth > 50) 

# Ethaboxam
mean_relative_growth %>% 
  filter(fungicide == "Ethaboxam" & dose == 100 & relative_growth > 50)

# remove insensitive isolates for analysis
no_EC50_data <- transform_fungicide %>% 
  filter(!(fungicide == "Azoxystrobin" & species == "Pythium vexans"))

no_EC50_data <- no_EC50_data 
```

```{r}
# check if there are isolates with hormetic effect
# >100% rel growth at lower dose
hormetic_isolates <- no_EC50_data %>% 
  group_by(fungicide, run, isolate_id, dose) %>% 
  summarise(relative_growth = mean(relative_growth)) %>% 
  filter(relative_growth > 100, 
         !dose == 0) %>% 
  dplyr::select(fungicide, run, isolate_id)

hormetic_isolates %>% 
  group_by(fungicide, run, isolate_id) %>% 
  n_distinct()

hormetic_data <- semi_join(no_EC50_data, hormetic_isolates, by = c("fungicide",
                                                                   "run", 
                                                                   "isolate_id"))
                                                                        
# when combine data from both runs
hormetic_isolates_both <- no_EC50_data %>% 
  group_by(fungicide, isolate_id, dose) %>% 
  summarise(relative_growth = mean(relative_growth)) %>% 
  filter(relative_growth > 100, 
         !dose == 0) %>% 
  dplyr::select(fungicide, isolate_id)

```

## fit log logistic models

```{r}
# fit log-logistic models with 3 and 4 parameters 
nest_fungicide <- no_EC50_data %>% 
  group_by(fungicide, isolate_id, species) %>% 
  nest()

fit_LL3 <- function(dataset){
  drm(relative_growth ~ dose, fct = LL.3(names = c("slope", "upper", "EC50")), data = dataset)
}

fit_LL4 <- function(dataset){
  drm(relative_growth ~ dose, fct = LL.4(names = c("slope", "lower", "upper", "EC50")), data = dataset)
}

log_models <- nest_fungicide %>% 
  mutate(ll3.model = purrr::map(data, fit_LL3), 
         ll4.model = purrr::map(data, fit_LL4), 
         ll3.fit = purrr::map(ll3.model, mselect), 
         ll4.fit = purrr::map(ll4.model, mselect) 
         )

log_selection <- log_models %>% 
  dplyr::select(ll3.fit, ll4.fit) %>% 
  gather(ll3.fit, ll4.fit, key = "model", value = "model.fit") %>%   
  unnest_wider(model.fit)

log_selection %>% 
  dplyr::select(fungicide, isolate_id, model, IC) %>% 
  spread(model, IC) %>% 
  filter(ll3.fit < ll4.fit) %>% 
  n_distinct() 

# lack of fit p-values 
log_selection %>% 
  filter(`Lack of fit` > 0.05) %>% 
  group_by(model)%>% 
  count() 
```

LL.3 model look better overall

```{r}
# extract absolute EC50s from LL.4 model
abs_EC50 <- function(models){
  ED(models, respLev = c(50), type = "absolute", interval = "delta")
}

log_models <- log_models %>% 
  mutate(ll3_EC50 = purrr::map(ll3.model, abs_EC50))

log_EC50 <- log_models %>% 
  dplyr::select(ll3_EC50 ) %>% 
  unnest_wider(ll3_EC50) %>% 
  rename(EC50 = ...1  , 
         std.error = ...2 ,
         lower = ...3, 
         upper = ...4)

# add isolates with no EC50
azo_iso <- mean_relative_growth %>% 
  filter(fungicide == "Azoxystrobin" & dose == 10 & relative_growth > 50)

etha_iso <- mean_relative_growth %>% 
  filter(fungicide == "Ethaboxam" & dose == 100 & relative_growth > 50)

# EC50 for plotting
no_EC50_isolates <- bind_rows(azo_iso, etha_iso, mef_met_iso) %>% 
  dplyr::select(isolate_id, species, fungicide) %>% 
  mutate(EC50 = case_when(
      fungicide == "Azoxystrobin" ~ 10, 
      fungicide == "Ethaboxam" ~ 100
    ))

# remove dose column
no_EC50_isolates <- no_EC50_isolates %>% 
  group_by(isolate_id, species, fungicide, EC50) %>% 
  ungroup() %>% 
  distinct(isolate_id, species, fungicide, EC50)

# add missing columns to combine with calculated EC50
no_EC50_isolates <- no_EC50_isolates %>% 
  mutate(std.error = NA, 
         lower = NA, 
         upper = NA)

# combine
all_EC50_data <- bind_rows(log_EC50, no_EC50_isolates )

# write_csv(all_EC50_data, here("output", "results", "03_fungicide-sensitivity_EC50.csv"))
```


## Figure 

```{r}
fungicide_EC50 <- read_csv(here("output", "results", "03_fungicide-sensitivity_EC50.csv"))

# reduce the binomials length, for plotting
fungicide_EC50 <- fungicide_EC50 %>% 
  mutate(short_species = case_when(
    species == "Pythium acrogynum" ~ "P. acrogynum", 
    species == "Pythium irregulare" ~ "P. irregulare", 
    species == "Pythium torulosum" ~ "P. torulosum",
    species == "Pythium ultimum var. ultimum" ~ "P. ultimum var. ultimum", 
    species == "Pythium yorkensis"  ~ "P. yorkensis", 
    species == "Pythium acanthophoron" ~ "P. acanthophoron", 
    species ==  "Pythium aphanidermatum" ~  "P. aphanidermatum", 
    species == "Pythium pleroticum"  ~ "P. pleroticum", 
    species == "Pythium ultimum var. sporangiiferum" ~ "P. ultimum var. sporangiiferum", 
    species == "Pythium vexans" ~ "P. vexans" , 
    TRUE ~ as.character(species)
    ))

figure5 <- ggplot(fungicide_EC50, aes(x = EC50, fill = short_species)) +
  geom_histogram(alpha = 0.5, color = "white", bins = 10) +
  facet_wrap(~ fungicide, scales = "free_x") +
  labs(x = expression('EC'[50]*' (μg/ml)'), 
       y = "Frequency", 
       fill = "") +
  theme(strip.text = element_text(face = "bold", size = rel(1)), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray95"), 
        panel.grid.major.y = element_line(color = "gray95"), 
        legend.position = "bottom") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))


ggsave("figure5.tiff", figure5, dpi = 300, 
       path = here("docs", "figures"), width = 5, height = 5)



```

```{r}

```


